Fermi and eROSITA bubbles

### 2021-12-13 paper.tex
update

\begin{abstract}
\end{abstract}

@@

\section{Methodology}
\subsection{Assumptions and Numerical Techniques}
  We used the GPU-accelerated special relativistic hydrodynamics AMR code (\textsc{gamer-sr}) developed at the\
  National Taiwan University\
  (Schive et al. \citeyear{gamer-1}, \citeyear{gamer-2}; \citeauthor{tseng2021} \citeyear{tseng2021}) to carry out\
  the simulations of Fermi and eROSITA bubbles. \textsc{gamer-sr} adopts a new algorithm \citep{tseng2021} to convert between\
  primitive and conserved variables, significantly reducing numerical error caused by catastrophic cancellations\
  that commonly occur within the regions with high Mach number. i.e. jet-ISM interaction zones.

  The simulation box is $14\times14\times28$ kpc, slightly larger than the size of eROISTA bubbles,\
  with a root grid $N=16\times16\times32$. We resolve the jet source located at the center of the simulation domain,\
  cells are refined to level 11, leading to a finest spatial resolution of 0.4 pc.\

  We treat CRs as a second fluid and solve directly for the evolution of CR pressure $p_{\text{cr}}$\
  as a function of $\mathbf{r}$ and $t$.\
  We did not model the CR energy spectrum and neglect the cooling and heating processes of CRs,\
  such as energy losses due to synchrotron and inverse Compton emission, and reacceleration in shocks.

@@

  \textsc{gamer-sr} solves the mass and energy-momentum conservation laws of special relativistic ideal fluid\
  including CR advection, and dynamical coupling between the thermal gas and CRs. The governing equations\
  can be written as

\begin{enumerate}
  \item Assumptions on hydrodynamics:
    \begin{enumerate}
      \item Ideal fluid with TM EoS (\citealt{Taub}; \citealt{TM_EOS}).
    \end{enumerate}
%  \item Assumptions on cosmic-ray:
%    \begin{enumerate}
%      \item We treat CRs as a second fluid and solve directly for the evolution of CR pressure $p_{\text{cr}}$\
%            as a function of $\mathbf{r}$ and $t$.
%      \item We did not model the CR energy spectrum.
%      \item We neglected the cooling and heating processes of CRs, such as energy losses due to synchrotron\
%            and inverse Compton emission, and reacceleration in shocks.
%      \item We have assumed cosmic-ray is passive. (i.e. $p_{\text{cr}}\ll p_{\text{gas}}$)
%      \item \textcolor{red}{We have ignored the $\mathbf{B}$ field within\
%            the simulation box as the field inside\
%            the bubbles should be weak due to adiabatic expansion, and thus the magnetic fields has\
%            little effect on the overall dynamics.\
%            We also assumed the magnetic field is highly entangled on small-scale (how small?), resulting in\
%            the negligible cosmic-ray diffusion.}
%    \end{enumerate}
  \item Assumptions on gravity:
    \begin{enumerate}
      \item We use the potential of isothermal slab, symmetric about $z=0$ and\
            supported by gas pressure and by under its own self-gravity,\
            to mimic the realistic potential of Galactic bulge.
      \item We still use Newtonian gravity, as the relativistic fluid ejected by the jet source\
            is quickly slowed down by heavy disk in a short time, and the relativistic fluid\
            accounts for a little minority of total mass inside the simulation box.
      \item The ISM disk and atmosphere are subjected to the fixed external potential\
            due to a disk bulge and dark matter halo.
      \item In addition to the gravitational interaction, we ignore other interactions between stars and gases.
      \item We also ignore the self-gravity of the ISM disk and of the atmosphere.
      \item We ignore the centrifugal force of Milky Way rotation acting on the bubbles.
      \item We use the potential of isothermal slab to mimic the gravitational potential due to the stellar bulge.
      \item The interface between cold ISM disk and atmosphere is parallel to the Galactic plane and is pressure balanced.
    \end{enumerate}
\end{enumerate}
We simulate 3D special relativistic hydrodynamics with passive CR injections from the GC using\
the special hydrodynamics GPU code \textsc{gamer-sr} \citep{tseng2021}.

@@

The CRs are advected with the thermal gas, but the gas cannot react to the CR pressure.\
In this approach, the CRs are treated as a single species without distinction between electrons and protons.\
We did not model the CR energy spectrum, and we neglected the cooling and heating processes of CRs,\
such as energy losses due to synchrotron and inverse Compton emission, and reacceleration in shocks.\

\begin{subequations}
  \label{conservative-form}
  \begin{align}
   &\partial_{t} D+\partial_{j} \left(DU^{j}/\gamma\right)=0,\label{D evolution}\\
   &\partial_{t} M^{i}+\partial_{j} \left(M^{i}U^{j}/\gamma+p_{\text{gas}}\delta^{ij}\right)=\
   -\rho\partial_{i}\Phi,\label{M evolution}\\
   &\partial_{t} \tilde{E}+\partial_j \left[\left(\tilde{E}+p_{\text{gas}}\right)U^{j}/\gamma\right]=0, \label{E evoltion}\\
   &\partial_{t} \left(\gamma e_{\text{cr}}\right) + \partial_{j} \left(e_{\text{cr}}U^{j}\right)=-p_{\text{cr}} \partial_{j} U^{j},\label{D evolution}
  \end{align}
\end{subequations}
where the five conserved quantities of gas $D$, $M^{i}$, and $\tilde{E}$ are the mass density,\
the momentum densities, and the reduced energy density, respectively.\
The reduced energy density is defined by subtracting the rest mass energy density of gas\
from the total energy density of gas.\
$\gamma$ and $U^{j}$ are the temporal and spatial component of four-velocity of gas.\
$\rho$ is the gas density in the local rest frame defined by $D/\gamma$.\
$p_{\text{gas}}$ is the gas pressure.\
$p_{\text{cr}}$ and $e_{\text{cr}}$ are the CR pressure and CR energy density measured in the local rest frame.\
$\Phi$ is the gravitation potential.\
$c$ is the speed of light, and $\delta^{ij}$ is the Kronecker delta notation.\
Throughout this paper, Latin indices run from 1 to 3, except when stated otherwise.

@@

Inverse Compton scattering
\begin{equation}
  \frac{dE}{dtd\epsilon_{1}dV} = \frac{3}{4}\sigma_{T}c\textit{C}\int^{\epsilon_{\text{max}}}_{\epsilon_{\text{min}}}
               \epsilon_{1}\frac{n(\epsilon)}{\epsilon}d\epsilon\int^{\gamma_{\text{max}}}_{\gamma_{\text{min}}}
               \gamma^{-(p+2)}f(q, \Gamma)d\gamma.
\end{equation}

\begin{equation}
f(q, \Gamma) = 2q\ln q+(1+2q)(1-q)+0.5(1-q)\frac{\left(\Gamma q\right)^2}{1+\Gamma q},
\end{equation}

@@

@@

\begin{equation}
\sigma(\epsilon_{\text{p}})=34.3+1.88L+0.25L^{2} \text{ mb}.
\end{equation}

@@

where $B=1.30+0.14L+0.011L^2$, $\beta=\left(1.79+0.11L+0.008L^{2}\right)^{-1}$,\
$\left(0.801+0.049L+0.014L^{2}\right)^{-1}$, $L=\ln(\epsilon_{\text{p}}/1 \text{ TeV})$.

>>>>

\begin{abstract}
\begin{enumerate}
  \item Release the caveat of jet direction.
  \item Rule out the hadronic process.
  \item Predict the small inner bubbles.
  \item How to explain why we assume e_CR << e_gas
  \item How to explain the lifespan of bubbles is too long.
  \item How to explain the jet power.
  \item How to explain the inconsistant of gamma-ray map.
\end{enumerate}
\end{abstract}

@@

\section{Methodology}

  We used the GPU-accelerated special relativistic hydrodynamics AMR code (\textsc{gamer-sr}) developed at the\
  National Taiwan University\
  (Schive et al. \citeyear{gamer-1}, \citeyear{gamer-2}; \citeauthor{tseng2021} \citeyear{tseng2021})\
  to carry out the simulations of the Fermi and eROSITA bubbles by CR and relativistic fluid injections from the GC.

  The CRs are advected with the thermal gas, and in return the velocities of gas\
  can react to the gradients of the CR pressure\
  via the presence of a source (or coupling) term containing spatial derivatives of fluid velocities.

  Since CR diuusion insignificantly affect on the overall mophology of the Fermi bubbles\
  but only sharpen the edges of the simulated bubbles by the interplay between anisotropic CR diffusion\
  and magnetic fields
with suppressed perpendicular diffusion across the bubble surface.

  Since the sharpness of bubble edges requires suppression of CR diffusion across the bubble surface
 has a insignificant effect on the overall mophology of the Fermi bubbles

  In this approach, the CRs are treated as a single species without distinction between electrons and protons.\
  We do not simulate the spectral evolution of the CR, and we neglected the cooling and heating processes of CRs,\
  such as energy losses due to synchrotron and inverse Compton emission, and reacceleration in shocks.\

  We treat CRs as a second fluid with a initially fixed energy\
  and solve directly for the evolution of CR pressure $p_{\text{cr}}$\
  as a function of $\mathbf{r}$ and $t$.\

@@

  We still use Newtonian gravity, as the relativistic fluid ejected by the jet source\
  is quickly slowed down by heavy disk in a short time, and the relativistic fluid\
  accounts for a little minority of total mass inside the simulation box.

  \textsc{gamer-sr} solves the mass and energy-momentum conservation laws of special relativistic ideal fluid\
  including CR advection, and dynamical coupling between the thermal gas and CRs without diffusion.\
  The governing equations can be written as

  \begin{subequations}
    \label{conservative-form}
    \begin{align}
     &\partial_{t} D+\partial_{j} \left(DU^{j}/\gamma\right)=0,\label{D evolution}\\
     &\partial_{t} M^{i}+\partial_{j} \left(M^{i}U^{j}/\gamma+p_{\text{gas}}\delta^{ij}\right)=\
     -\rho\partial_{i}\Phi,\label{M evolution}\\
     &\partial_{t} \tilde{E}+\partial_j \left[\left(\tilde{E}+p_{\text{gas}}\right)U^{j}/\gamma\right]=0, \label{E evoltion}\\
     &\partial_{t} \left(\gamma e_{\text{cr}}\right) + \partial_{j} \left(e_{\text{cr}}U^{j}\right)=\
     -p_{\text{cr}} \partial_{j} U^{j},\label{D evolution}
    \end{align}
  \end{subequations}

  where the five conserved quantities of gas $D$, $M^{i}$, and $\tilde{E}$ are the mass density,\
  the momentum densities, and the reduced energy density, respectively.\
  The reduced energy density is defined by subtracting the rest mass energy density of gas\
  from the total energy density of gas.\
  $\gamma$ and $U^{j}$ are the temporal and spatial component of four-velocity of gas.\
  $\rho$ is the gas density in the local rest frame defined by $D/\gamma$.\
  $p_{\text{gas}}$ is the gas pressure.\
  $p_{\text{cr}}$ and $e_{\text{cr}}$ are the CR pressure and CR energy density measured in the local rest frame.\
  $\Phi$ is the gravitation potential.\
  $c$ is the speed of light, and $\delta^{ij}$ is the Kronecker delta notation.\
  Throughout this paper, Latin indices run from 1 to 3, except when stated otherwise.

  \textsc{gamer-sr} adopts a new algorithm \citep{tseng2021} to convert between\
  primitive ($p$, $\rho$, $U^{j}$) and conserved variables, significantly reducing numerical error\
  caused by catastrophic cancellations\
  that commonly occur within the regions with high Mach number. e.g., jet-ISM interaction zones.

\subsection{Initial conditions for the Fermi and eROSITA bubbles}
  The simulation box is $14\times14\times28$ kpc, slightly larger than the size of eROISTA bubbles,\
  with a root grid $N=16\times16\times32$. We resolve the jet source located at the center of the simulation domain,\
  cells are refined to level 11, leading to a finest spatial resolution of 0.4 pc.\

\begin{enumerate}
  \item Assumptions on hydrodynamics:
    \begin{enumerate}
      \item Ideal fluid with TM EoS (\citealt{Taub}; \citealt{TM_EOS}).
    \end{enumerate}
%  \item Assumptions on cosmic-ray:
%    \begin{enumerate}
%      \item We treat CRs as a second fluid and solve directly for the evolution of CR pressure $p_{\text{cr}}$\
%            as a function of $\mathbf{r}$ and $t$.
%      \item We did not model the CR energy spectrum.
%      \item We neglected the cooling and heating processes of CRs, such as energy losses due to synchrotron\
%            and inverse Compton emission, and reacceleration in shocks.
%      \item We have assumed cosmic-ray is passive. (i.e. $p_{\text{cr}}\ll p_{\text{gas}}$)
%      \item \textcolor{red}{We have ignored the $\mathbf{B}$ field within\
%            the simulation box as the field inside\
%            the bubbles should be weak due to adiabatic expansion, and thus the magnetic fields has\
%            little effect on the overall dynamics.\
%            We also assumed the magnetic field is highly entangled on small-scale (how small?), resulting in\
%            the negligible cosmic-ray diffusion.}
%    \end{enumerate}
  \item Assumptions on gravity:
    \begin{enumerate}
      \item We use the potential of isothermal slab, symmetric about $z=0$ and\
            supported by gas pressure and by under its own self-gravity,\
            to mimic the realistic potential of Galactic bulge.
%      \item We still use Newtonian gravity, as the relativistic fluid ejected by the jet source\
%            is quickly slowed down by heavy disk in a short time, and the relativistic fluid\
%            accounts for a little minority of total mass inside the simulation box.
      \item The ISM disk and atmosphere are subjected to the fixed external potential\
            due to a disk bulge and dark matter halo.
      \item In addition to the gravitational interaction, we ignore other interactions between stars and gases.
      \item We also ignore the self-gravity of the ISM disk and of the atmosphere.
      \item We ignore the centrifugal force of Milky Way rotation acting on the bubbles.
      \item We use the potential of isothermal slab to mimic the gravitational potential due to the stellar bulge.
      \item The interface between cold ISM disk and atmosphere is parallel to the Galactic plane and is pressure balanced.
    \end{enumerate}
\end{enumerate}
We simulate 3D special relativistic hydrodynamics with passive CR injections from the GC using\
the special hydrodynamics GPU code \textsc{gamer-sr} \citep{tseng2021}.

@@

@@

Inverse Compton scattering
\begin{equation}
  \frac{dE}{dtd\epsilon_{1}dV} = \frac{3}{4}\sigma_{T}c\textit{C}\epsilon_{1}\int^{\epsilon_{\text{max}}}_{\epsilon_{\text{min}}}
               \frac{n(\epsilon)}{\epsilon}d\epsilon\int^{\gamma_{\text{max}}}_{\gamma_{\text{min}}\left(\epsilon\right)}
               \gamma^{-(p+2)}f(q, \Gamma)d\gamma.
\end{equation}

\begin{equation}
f(\gamma_{\text{e}}, \epsilon) = 2q\ln q+(1+2q)(1-q)+0.5(1-q)\frac{\left(\Gamma q\right)^2}{1+\Gamma q},
\end{equation}

@@

$\gamma_{\text{min}}\left(\epsilon\right)$ is the root of $f(\gamma_{\text{e}}, \epsilon)=0$ at a specific incident photon energy.

@@

\begin{equation}
\sigma(\epsilon_{\text{p}})=34.3+1.88L+0.25L^{2}\left[1-\frac{E_{\text{threshold}}}{E_{\text{p}}}\right]^{4} \text{ mb}.
\end{equation}

@@

where
$x=\epsilon_{1}/\epsilon_{\text{p}}$, $B=1.30+0.14L+0.011L^2$, $\beta=\left(1.79+0.11L+0.008L^{2}\right)^{-1}$,\
$\left(0.801+0.049L+0.014L^{2}\right)^{-1}$, $L=\ln(\epsilon_{\text{p}}/1 \text{ TeV})$.
